\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)

### 2021-11-25 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Bricoles sur les effets de composition
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

Going into a more detailed analysis by transport mode, we do not find much evidence of a substantial trend difference for air transport between the raw transport costs measure and the pure transport costs until the mid-2000s. Inspecting the panels (a) to (c) of Figure \ref{fig:totalTC_compeffects_excl}, for all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely until around 2005. Afterwards however, composition effects are the main driver of the reduction in the raw series, while pure transport costs are roughly constant, especially for the additive component.

Composition effects do also play a role in vessel transport, yet limited in magnitude. For all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)), they amplify the reduction in ``pure'' transport costs at the beginning of the period in the 1970s (the dotted line is below the plain line, and the gap is increasing). Afterwards, the gap between the two remains roughly constant across years, indicating that the amplification effect does not strengthen anymore: the two trends are similar.
Considering the raw series (dotted line), vessel transport costs have decreased by 60\% over the period, which can be decomposed into a 50\% decrease in pure transport costs (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two). This is particularly the case for the multiplicative component (panel (d)).\smallskip

>>>>

Going into a more detailed analysis by transport mode, we do not find much evidence of a substantial trend difference for air transport between the raw transport costs measure and the pure transport costs until the mid-2000s: Inspecting the panels (a) to (c) of Figure \ref{fig:totalTC_compeffects_excl}, for all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line stay close together until around 2005. Afterwards however, composition effects are the main driver of the reduction in the raw series, while pure transport costs are roughly constant, especially for the additive component.

Composition effects do also play a role in vessel transport. It is limited in magnitude. For all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)), they amplify the reduction in ``pure'' transport costs in the 1970s (the dotted line is below the plain line, and the gap is increasing). Afterwards, the gap between the two remains roughly constant across years, indicating that the amplification effect does not strengthen anymore: the two trends are similar.
Considering the raw series (dotted line), vessel transport costs have decreased by 60\% over the period, which can be decomposed into a 50\% decrease in pure transport costs (dotted line), and a 10\% reduction that comes from composition effects (the difference between the two line). This is particularly the case for the multiplicative component (panel (d)).\smallskip

### 2021-11-25 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
Bricoles sur les effets de composition
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

In this section, we refine the characterization of the evolution of international transport costs by distinguishing primary goods trade flows and and manufactured goods trade flows. The evolution in transport costs over time, by transport mode (overall transport costs and composition effects excluded) are reported in Figure \ref{fig:totalTC_compeffects_excl_manuf} for the manufacturing sector, and in Figure \ref{fig:totalTC_compeffects_excl_primary} for the primary sector. To ease the comparison, we also report the results obtained on the whole range of trade flows (i.e., Figure 2 of the paper), in Figure \ref{fig:totalTC_compeffects_excl}.

The classification retained to categorize trade flows follows the UNCTAD classification (on STIC Revision 3)\footnote{See "UNCTAD product groupings and composition (SITC Rev. 3)" in http://unctadstat.unctad.org/EN/Classifications.html, accessed Septembre 2018}.
Are considered as ``primary goods'' all flows recorded as ``Food and live animals'' (First digit ``0'' in the SITC Classification), ``Beverages and tobacco'' (First digit ``1''), ``Crude materials, inedible, except fuels'' (First digit ``2''), ``Mineral fuels, lubricants and related materials'' (First digit ``3''), ``Animal and vegetable oils, fats and waxes'' (First digit ``4''), ``Pearls, precious \& semi-precious stones'' (Classified ``667'' in the SITC Classification) and ``Non-ferrous metals'' (classified ``68'' in the SITC Classification).

@@

As reported in Figures \ref{fig:totalTC_compeffects_excl_manuf} and \ref{fig:totalTC_compeffects_excl_primary}, both the ``ceteris paribus'' transport costs and the raw measure have regularly declined over the period in both sectors, by roughly the same order of magnitude (50\% in air, 60\% in vessel for overall transport costs, panels (c) and (f)). However the role of trade composition effects in accounting for this trend pattern differs depending on the sector.\\
In the manufacturing sector, Figure \ref{fig:totalTC_compeffects_excl_manuf} reports a very similar time trend decomposition than what is obtained on the whole range of goods (Figure \ref{fig:totalTC_compeffects_excl}).  One can further notice that in vessel transport, composition effects play virtually no role at the end of the period, as most of the decrease can be imputed to the reduction of ``ceteris paribus'' transport costs (the blue continuous line)

The situation is different in the primary sector. In air transport, the role of composition effects exhibits an irregular pattern over time. They dampen the reduction in pure transport costs up to the 2000s. Afterwards, they conversely more than compensate the increase in pure transport costs (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (a), (b) and (c)). In vessel transport, the composition effects play a limited role in the primary sector as for the whole range of goods matter (Figure \ref{fig:totalTC_compeffects_excl_primary}, right-hand panels (d), (e) and (f)), by slightly amplifying the reduction in the pure transport costs in both cases.

One explanation for the similarity between the results for the manufacturing goods trade and for total trade can be found in the share of primary goods in total flows as reported in Figure \ref{fig:Share_prim_goods}. In air transport, the share of primary goods in the total value of US imports is very small, around 10\%. Primary goods make a higher proportion of trade flows in vessel transport, especially over 1974-1982 (between 40\% and 60\%). On the following sub-period though, their share has fallen to 20-30\%. Given the modest proportion of primary goods in total import flows of the US economy compared to the manufactured sector, it is not surprising that the diagnosis made about the time trend of transport costs when all types of flows are considered is driven by the manufacturing sector.

>>>>

In this section, we refine the characterization of the evolution of international transport costs by distinguishing primary and manufactured goods trade flows.
The evolution in transport costs over time, by transport mode (overall transport costs and composition effects excluded) is reported in Figure \ref{fig:totalTC_compeffects_excl_manuf} for the manufacturing sector, and in Figure \ref{fig:totalTC_compeffects_excl_primary} for the primary sector. To ease the comparison, we also report the results obtained on the whole range of trade flows (i.e., Figure 2 of the paper), in Figure \ref{fig:totalTC_compeffects_excl}.

The classification retained to categorize trade flows follows the UNCTAD classification (on STIC Revision 3)\footnote{See "UNCTAD product groupings and composition (SITC Rev. 3)" in http://unctadstat.unctad.org/EN/Classifications.html, accessed Septembre 2018}.
``Primary goods'' flows are recorded as ``Food and live animals'' (First digit ``0'' in the SITC Classification), ``Beverages and tobacco'' (First digit ``1''), ``Crude materials, inedible, except fuels'' (First digit ``2''), ``Mineral fuels, lubricants and related materials'' (First digit ``3''), ``Animal and vegetable oils, fats and waxes'' (First digit ``4''), ``Pearls, precious \& semi-precious stones'' (Classified ``667'' in the SITC Classification) and ``Non-ferrous metals'' (classified ``68'' in the SITC Classification).
All the other flows are part of ``Manufacturing''.

@@

As reported in Figures \ref{fig:totalTC_compeffects_excl_manuf} and \ref{fig:totalTC_compeffects_excl_primary}, both the ``pure'' transport costs and the raw measure have regularly declined over the period in both sectors, by roughly the same order of magnitude (50\% in air, 60\% in vessel for overall transport costs, panels (c) and (f)). However the role of trade composition effects in accounting for this trend depends on the sector.\\
In the manufacturing sector, Figure \ref{fig:totalTC_compeffects_excl_manuf} reports a very similar time trend decomposition than what is obtained on the whole range of goods (Figure \ref{fig:totalTC_compeffects_excl}).  One can further notice that in vessel transport, composition effects play virtually no role at the end of the period, as most of the decrease can be imputed to the reduction of ``pure'' transport costs (the blue continuous line)

The situation is different in the primary sector. In air transport, the role of composition effects exhibits an irregular pattern over time. They dampen the reduction in pure transport costs up to the 2000s. Afterwards, they conversely more than compensate the increase in pure transport costs (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (a), (b) and (c)). In vessel transport, the composition effects play a limited role in the primary sector, similarly to their role for the whole range of goods (Figure \ref{fig:totalTC_compeffects_excl_primary}, right-hand panels (d), (e) and (f)), by slightly amplifying the reduction in the pure transport costs.

One explanation for the similarity between the results for the manufacturing goods trade and for total trade can be found in the share of primary goods in total flows as reported in Figure \ref{fig:Share_prim_goods}. In air transport, the share of primary goods in the total value of US imports is very small, around 10\%. Primary goods make a higher proportion of trade flows in vessel transport, especially over 1974-1982 (between 40\% and 60\%). On the following sub-period though, their share has fallen to 20-30\%. Given the modest proportion of primary goods in total import flows of the US economy compared to the manufactured goods, it is not surprising that the diagnosis made about the time trend of transport costs when all types of flows are considered is driven by the manufacturing.
